Effects of rotation on the Lyman-$\alpha$ line morphology in distant galaxies

### 2013-02-22 paper/Analisis.tex
new text edits, issues from this version were already posted

\newcommand{\LCDM}{$\Lambda$CDM~}
\newcommand{\beq}{\begin{eqnarray}}  
\newcommand{\eeq}{\end{eqnarray}}  
\newcommand{\zz}{$z\sim 3$} 
\newcommand{\apj}{ApJ}  
\newcommand{\apjs}{ApJS}  
\newcommand{\apjl}{ApJL}  
\newcommand{\aj}{AJ}  
\newcommand{\mnras}{MNRAS}  
\newcommand{\mnrassub}{MNRAS accepted}  
\newcommand{\aap}{A\&A}  
\newcommand{\aaps}{A\&AS}  
\newcommand{\araa}{ARA\&A}  
\newcommand{\nat}{Nature}  
\newcommand{\physrep}{PhR}
\newcommand{\pasp}{PASP}    
\newcommand{\pasj}{PASJ}    
\newcommand{\avg}[1]{\langle{#1}\rangle}  
\newcommand{\ly}{{\ifmmode{{\rm Ly}\alpha}\else{Ly$\alpha$}\fi}}
\newcommand{\hMpc}{{\ifmmode{h^{-1}{\rm Mpc}}\else{$h^{-1}$Mpc }\fi}}  
\newcommand{\hGpc}{{\ifmmode{h^{-1}{\rm Gpc}}\else{$h^{-1}$Gpc }\fi}}  
\newcommand{\hmpc}{{\ifmmode{h^{-1}{\rm Mpc}}\else{$h^{-1}$Mpc }\fi}}  
\newcommand{\hkpc}{{\ifmmode{h^{-1}{\rm kpc}}\else{$h^{-1}$kpc }\fi}}  
\newcommand{\hMsun}{{\ifmmode{h^{-1}{\rm {M_{\odot}}}}\else{$h^{-1}{\rm{M_{\odot}}}$}\fi}}  
\newcommand{\hmsun}{{\ifmmode{h^{-1}{\rm {M_{\odot}}}}\else{$h^{-1}{\rm{M_{\odot}}}$}\fi}}  
\newcommand{\Msun}{{\ifmmode{{\rm {M_{\odot}}}}\else{${\rm{M_{\odot}}}$}\fi}}  
\newcommand{\msun}{{\ifmmode{{\rm {M_{\odot}}}}\else{${\rm{M_{\odot}}}$}\fi}}  
\newcommand{\lya}{{Lyman$\alpha$~}}
\newcommand{\clara}{{\texttt{CLARA}}~}
\newcommand{\rand}{{\ifmmode{{\mathcal{R}}}\else{${\mathcal{R}}$ }\fi}}  
\newcommand{\hs}{{\hspace{1mm}}}

@@

\title[Rotation in the Lyman-$\alpha$ line]{Effects of bulk gas rotation on
  the emergent Lyman-$\alpha$ line in distant galaxies}
\author[N. Garavito and J.E. Forero-Romero]{
\parbox[t]{\textwidth}{\raggedright 
  Nicolas Garavito-Camargo.$^{1}$ 
  Jaime E. Forero-Romero$^{2}$ 
}
\vspace*{6pt}\\
$^{1}$Uni A
$^{2}$Uni B
}
\maketitle

@@

Due to the resonant nature of the lyman alpha line, gas kinematics
play an important role in the shape of this line. In particular we
study the case in which this gas is spherically symmetric and its
rotating.

@@

%In cartesian coordinates the rotational velocity around de the z-axys
%is defined by:

@@

%Where $x$ and $y$ define de position of the gas atoms and dust
%particles. 
%The minus sign in the x-component of the velocity is due to
%the direction of rotation, in this case we assume that as seeing from
%the top,the rotation is anticlockwise.
where $R$ is the radius of the sphere and $V_{\rm max}$ is the linear
velocity at the sphere's surface. The minus sign in the x-component of the velocity is due to
the direction of rotation, in this case we assume that the angular
velocity vector goes in the $\hat{k}$ direction.

@@

%We propose several models which include differents rotational
%velocities and optical depths. We also take into account the dust
%within the gas and two differents initial distributions of the
%photons, central and homogeneous. At the end we study 48 models, those
%are resumed in Table[1]\\

@@

% JF: digo que son 60 modelos porque estoy incluyendo los de velocidad cero.

%We later study this models taking into account the posititon of the
%observer respect to the axis of rotation of the cloud.

Additionaly in the postprocessing process we will take into account
different observer positions with respect to the cloud rotation
axis.

\begin{table}
\begin{center}
\begin{tabular}{ccc}\hline
Velocity (km s$^{-1}$) & $V_{\rm max}$&$0,\ 50,\ 100,\ 200,\ 300$\\
Hydrogen Optical Depth & $\tau_{H} $ & $10^{5},\ 10^{6},\ 10^{7}$\\
Dust Optical Depth & $\tau_{A}$ & XXX \\
Photons Distributions & & Central, Homogeneous\\
\hline
\end{tabular}
\caption{
Values for the varying input parameters in CLARA. Taking into account
all the possible combinations for these models
} 
\label{table:models}
\end{center}
\end{table}

\begin{figure*}
  \includegraphics[width=0.45\textwidth]{7tDifSpeedsZ.png}
  \includegraphics[width=0.45\textwidth]{7tHOMDifSpeeds1.png}
  \label{fig:HOMandCentral}\caption{Shape of the lyman alpha line for
    differents velocities. The left panel show the central
    distribution while the right panel show the homogeneous
    distribution.}
\end{figure*}

Now if we take into account the position of the observer and compute
this for differents angles of observation $\theta$ the outcoming
spectra is modified as is shown in Figure~\ref{DifferentObservers}. The main feature here is that as the angle increases the line height decreases. It means that a observer in the poles will observe a higher double peak than a
observer in the ecuator. We will understand this in more detail when
we present the results of the scape fraction in section~\ref{sec:EF}.

\begin{figure}
\begin{center}
\includegraphics[scale=0.45]{Observers.png}
\end{center}
\caption{Spectra for different observers. Model:$V=300km/s$, Optical
  Depth $\tau=10^{7}$ and Central Distribution without dust.} 
  \label{DifferentObservers}
\end{figure}

%describir los modelos, poner las ecuaciones de la velocidad de
%rotacion, %poner las graficas para un modelo dado poner el efecto de
%la rotación (poner %uno con velocidad cero) y del observador.

\section{Results}
\label{sec:Results}

%Ver correo de Jaime y tratar de explicar a que se deben estos efectos.

In the previous section we describe the models that we use to study the
effect of rotation in the Lyman alpha line for this we saw the outcoming specrtums including rotation. Now we show the main results that we
obtain, in particual we pay spetial attention to the position of the
maximums, the width of the line and the escape fraction in function of
the rotational velocity and the angle of observation $\theta$.

In Figure~\ref{fig:HOMandCentral} the effect that rotation have in the spectra is shown,
for both distributions central (Left) and homogeneous (right). This is
a general case in the sense that is independent of the position of an
external observer.

@@

The maximums position gives information about the wave lenght of the
mayority of the outcoming photons after they interact with the gas, in
addition as a photon have more scatterings its wave lenght would be
larger than the initial which is $1216$ {\AA}. 
As we can see in Figure ~\ref{figure:maximumsvsvelocities} the position of the maximums does not change with rotational velocity for the central dsitribution. On the other hand for the homogeneous distriution the maxima position $X_{m}$ change for $Log_{10}\tau=5$, this is because at higher rotational velocities photons escape with less scatterings, for $v=100\dfrac{km}{s}$ the majority of the photons still escape with scatterings but for $v=200\dfrac{km}{s}$ the number of photons that escape with a few or any scatterings is higher, this is why the maxima change to the center of the spectra.\\

@@

% fix this table is too big
\begin{table}
\begin{center}
\begin{tabular}{cc}\hline
Model & Fit\\
Central, $\tau=10^5, V=200\dfrac{km}{s}$ &$EW(\mu)=-59-1677\mu-17864\mu^{2}+84509\mu^{3}-149877\mu^{4}$\\
Central, $\tau=10^6, V=200\dfrac{km}{s}$ &$EW(\mu)=-1+53.8\mu-1080\mu^{2}+9644\mu^{3}-32264\mu^{4}$\\
Central, $\tau=10^7, V=200\dfrac{km}{s}$ & $EW(\mu)=0.011+0.9\mu-27\mu^{2}+359\mu^{3}-1791\mu^{4}$\\
Hom, $\tau=10^5, V=200\dfrac{km}{s}$ & $EW(\mu)=11.66-187.22\mu-54.58\mu^{2}+12932.5\mu^{3}-51882.9\mu^{4}$ \\
Hom, $\tau=10^6, V=200\dfrac{km}{s}$ & $EW(\mu)=-361.64+17158.92\mu-305268.4\mu^{2}+2413478\mu^{3}-7154649\mu^{4}$ \\
Hom, $\tau=10^7, V=200\dfrac{km}{s}$ & $EW(\mu)=-0.28+23.29\mu-721\mu^{2}+9912\mu^{3}-51074\mu^{4}$ \\
\hline
\end{tabular}
\caption{Fits for EW models} 
\label{table:fits}
\end{center}
\end{table}

@@

>>>>

\newcommand{\LCDM}{$\Lambda$CDM~}
\newcommand{\beq}{\begin{eqnarray}}  
\newcommand{\eeq}{\end{eqnarray}}  
\newcommand{\zz}{$z\sim 3$} 
\newcommand{\apj}{ApJ}  
\newcommand{\apjs}{ApJS}  
\newcommand{\apjl}{ApJL}  
\newcommand{\aj}{AJ}  
\newcommand{\mnras}{MNRAS}  
\newcommand{\mnrassub}{MNRAS accepted}  
\newcommand{\aap}{A\&A}  
\newcommand{\aaps}{A\&AS}  
\newcommand{\araa}{ARA\&A}  
\newcommand{\nat}{Nature}  
\newcommand{\physrep}{PhR}
\newcommand{\pasp}{PASP}    
\newcommand{\pasj}{PASJ}    
\newcommand{\avg}[1]{\langle{#1}\rangle}  
\newcommand{\ly}{{\ifmmode{{\rm Ly}\alpha}\else{Ly$\alpha$}\fi}}
\newcommand{\hMpc}{{\ifmmode{h^{-1}{\rm Mpc}}\else{$h^{-1}$Mpc }\fi}}  
\newcommand{\hGpc}{{\ifmmode{h^{-1}{\rm Gpc}}\else{$h^{-1}$Gpc }\fi}}  
\newcommand{\hmpc}{{\ifmmode{h^{-1}{\rm Mpc}}\else{$h^{-1}$Mpc }\fi}}  
\newcommand{\hkpc}{{\ifmmode{h^{-1}{\rm kpc}}\else{$h^{-1}$kpc }\fi}}  
\newcommand{\hMsun}{{\ifmmode{h^{-1}{\rm {M_{\odot}}}}\else{$h^{-1}{\rm{M_{\odot}}}$}\fi}}  
\newcommand{\hmsun}{{\ifmmode{h^{-1}{\rm {M_{\odot}}}}\else{$h^{-1}{\rm{M_{\odot}}}$}\fi}}  
\newcommand{\Msun}{{\ifmmode{{\rm {M_{\odot}}}}\else{${\rm{M_{\odot}}}$}\fi}}  
\newcommand{\msun}{{\ifmmode{{\rm {M_{\odot}}}}\else{${\rm{M_{\odot}}}$}\fi}}  
\newcommand{\lya}{{Lyman$\alpha$~}}
\newcommand{\clara}{{\texttt{CLARA}}~}
\newcommand{\rand}{{\ifmmode{{\mathcal{R}}}\else{${\mathcal{R}}$ }\fi}}  
\newcommand{\hs}{{\hspace{1mm}}}  
\newcommand{\kms}{\,km~s$^{-1}$}

@@

\title[Rotation in the Lyman-$\alpha$ line]{Effects of rotation on the
  Lyman-$\alpha$ line morphology in distant galaxies}
\author[N. Garavito and J.E. Forero-Romero]{
\parbox[t]{\textwidth}{\raggedright 
  Nicolas Garavito-Camargo.$^{1}$ 
  Jaime E. Forero-Romero$^{2}$ 
}
\vspace*{6pt}\\
$^{1}$Uni A
$^{2}$Uni B
}
\maketitle

@@

Due to the resonant nature of the lyman alpha line, gas kinematics
play an important role shaping its morphology. In the literature there
has been extensive studies of outflow/inflow configurations.

In this paper we study for the first time the impact of rotation on
the morphology of the Lyman $\alpha$ line. To isolate the effects of
rotation we focus on a simple system: the gas distribution is
spherical, with homogenous density and the gas rotates as a solid
body.

This paper is paper is structured as follows.

@@

@@

where $R$ is the radius of the sphere and $V_{\rm max}$ is the linear
velocity at the sphere's surface. The minus sign in the x-component of
the velocity indicates the direction of rotation, in this case we
assume that the angular velocity vector goes in the $\hat{k}$
direction.  The linear dependence of the velocity on the radial
distance describes the case of constant angular velocity
$\omega=V_{\rm max}/R$.

We take the polar angle $\theta$ that a unitary vector makes with the
rotation axis as defined by the dot product $\cos\theta =
{\hat{u}\cdot\hat{k}}$. In the Section \ref{sec:results} we will
present in detail how the line differs at different observing angles
$\theta$.

@@

@@

\begin{table}
\begin{center}
\begin{tabular}{ccc}\hline
Velocity (\kms) & $V_{\rm max}$&$0,\ 50,\ 100,\ 200,\ 300$\\
Hydrogen Optical Depth & $\tau_{H} $ & $10^{5},\ 10^{6},\ 10^{7}$\\
Dust Optical Depth & $\tau_{A}$ & XXX \\
Photons Distributions & & Central, Homogeneous\\
\hline
\end{tabular}
\caption{
Values for the varying input parameters in CLARA. Taking into account
all the possible combinations for these models
} 
\label{table:models}
\end{center}
\end{table}

%poner las graficas para un modelo dado poner el efecto de
%la rotación (poner %uno con velocidad cero) y del observador.

\section{Results}
\label{sec:results}

The central result of this paper is summarized in Figure
\ref{fig:HOMandCentral} where we show that rotation as a considerable
effect on the morphology of the emergent Lyman-alpha line both in the
case where the photons are emitted at the sphere's center and when
they are initialized with an homogeneous distribution all over the
gas volume.

The results for this outgoing spectra are integrated over the whole
sphere, meaning that all the escaping photons were taking into account
regardless of the direction of the outgoing photons. Figure
\ref{fig:Different Observers} shows how if one gives a weight to each outgoing
photon according to its direction when escaping the gas distribution
it is posible to detect notable differences in the spectrum for
different viewing angles.

\begin{figure*}
  \includegraphics[width=0.45\textwidth]{7tDifSpeedsZ.png}
  \includegraphics[width=0.45\textwidth]{7tHOMDifSpeeds1.png}
  \label{fig:HOMandCentral}\caption{Shape of the lyman alpha line for
    differents velocities. The left (right) panel shows the central
    (homogeneous) photon distribution.}
\end{figure*}

\begin{figure}
\begin{center}
\includegraphics[scale=0.45]{Observers.png}
\end{center}
\caption{Spectra for different observers. Model:$V=300$\kms, Optical
  Depth $\tau=10^{7}$ and Central Distribution without dust.} 
  \label{fig:DifferentObservers}
\end{figure}

@@

The maximums position gives information about the wave lenght of the
mayority of the outcoming photons after they interact with the gas, in
addition as a photon have more scatterings its wave lenght would be
larger than the initial which is $1216$ {\AA}. 
As we can see in Figure ~\ref{figure:maximumsvsvelocities} the
position of the maxima does not change with rotational velocity for
the central dsitribution. On the other hand for the homogeneous
distriution the maxima position $X_{m}$ change for $Log_{10}\tau=5$,
this is because at higher rotational velocities photons escape with
less scatterings, for $v=100$\kms the majority of the photons
still escape with scatterings but for $v=200$\kms the number
of photons that escape with a few or any scatterings is higher, this
is why the maxima change to the center of the spectra.\\

@@

% fix this table is too big
\begin{table*}
\begin{center}
\begin{tabular}{cc}\hline
Model & Fit\\
Central, $\tau=10^5, V=200$\kms &$EW(\mu)=-59-1677\mu-17864\mu^{2}+84509\mu^{3}-149877\mu^{4}$\\
Central, $\tau=10^6, V=200$\kms &$EW(\mu)=-1+53.8\mu-1080\mu^{2}+9644\mu^{3}-32264\mu^{4}$\\
Central, $\tau=10^7, V=200$\kms & $EW(\mu)=0.011+0.9\mu-27\mu^{2}+359\mu^{3}-1791\mu^{4}$\\
Hom, $\tau=10^5, V=200$\kms & $EW(\mu)=11.66-187.22\mu-54.58\mu^{2}+12932.5\mu^{3}-51882.9\mu^{4}$ \\
Hom, $\tau=10^6, V=200$\kms & $EW(\mu)=-361.64+17158.92\mu-305268.4\mu^{2}+2413478\mu^{3}-7154649\mu^{4}$ \\
Hom, $\tau=10^7, V=200$\kms & $EW(\mu)=-0.28+23.29\mu-721\mu^{2}+9912\mu^{3}-51074\mu^{4}$ \\
\hline
\end{tabular}
\caption{Fits for EW models} 
\label{table:fits}
\end{center}
\end{table*}

@@

... The results derived in this paper have consequences on the
interpreation of galaxy observations in the Lyman alpha line.
